\texorpdfstring{Identifying Software Performance Changes Across\\ Variants and Versions}{Identifying Software Performance Changes Across Variants and Versions}

### 2020-08-26 paper.tex
discussion changes and added doi's

	\section{Discussion}\label{sec:discussion}
	\paragraph{Efficiency} Among all experiments conducted with our synthetic setup, our approach yielded reasonable to high accuracy across most parameter settings. A subsequent analysis of parameter settings showed that our algorithm is not limited by the size of a problem space, but rather the number of measurements per iteration. 
	% mehr Interpretation, Details ausblenden
	Our algorithm found multiple change points across settings of up to 2,500 commits and and 2$^{64}$ configurations, resulting in a search space of about $4.6 \cdot 10^{22}$ possible measurements.
	In the context of not more than 5,000 measurements per iteration, the maximum number of measurements (after 30 iterations) added to the initial set of measurements is 150,000. For a setting with 2,500 commits, this represents around $1.4 \cdot 10^{-6}$\,\% and $3.2\cdot 10^{-16}$\,\% of possible measurements for settings with 32 and 64 options, respectively. The vast majority of cases terminated in early iterations. Hence, we are confident that our approach scales and can efficiently approximate change points for large software systems. Notably, we are able to reproduce this finding in real-world settings when identifying the temporal location of change points. For our controlled experiment setup, we have not tuned the algorithms’ parameters to reach optimal efficiency or accuracy. However, even without tuning, we demonstrated the feasibility and efficiency of our approach. Project-specific fine tuning of several parameters is a possible extension to our approach and can further improve our results.

	\paragraph{Pinpointing Configuration Options}
	The synthetic setup has shown that our algorithm can conceptually handle both time and space. We were able to precisely identify commits for which performance changed substantially. By contrast, we were not able to fully reproduce this in a setting with real-world data when attempting to pinpoint change points to configuration options. Possible explanations in the case of \textsc{lrzip} include a partly inconsistent configuration interface. 
	The reason for that is not a limitation of our approach, but missing information in the real-world data (which makes the synthetic setup so important). So, we have documented that, given a precisely identified commit, we can link the affected files to configuration options with little effort. In turn, given a set of code segments of a commit, matching this with the overall code base has been extensively studied before under the umbrella of feature location~\cite{hill_which_2013}.
	Inconsistencies in the documentation of configuration options is a well known and prevalent problem~\cite{rabkin_static_2011}. Our approach might be a means to complement existing feature location techniques with information about configuration-dependent performance changes. 	
	In our latter experiment, instead of directly sampling new arbitrary configurations, we used a finite set of configurations as a proxy (cf.~Section~\ref{sec:rq2_operationalization}). This does not reflect the original layout of our algorithm, but lets us efficiently explore some limitations of this approach. We have learned that, for interpreting the results of our approach, domain knowledge is advantageous as we have identified incomplete and inconsistent configuration interfaces as possible limiting factors. We acknowledge that, while our approach performs well as an automated pipeline in a controlled environment, for real-world applications, a semi-automated setup with additional manual measurement prioritization and deeper knowledge of configuration constraints would improve the applicability of our approach. This is a promising further avenue of future work.

@@

	\section{Conclusion}
	Changes in the observed performance of a software system can be configuration-specific and difficult to pinpoint to configuration options and commits. We combine both two orthogonal perspectives, the commit history and configurability of software systems to pinpoint performance changes to individual commits and configuration options. 
	We propose an iterative sampling approach that leverages similarities in performance histories of related configurations to identify performance shifts precisely and, subsequently, pinpoint them to configuration options and interactions. 
	We have shown with a synthetic data set that our approach is sound and able to identify performance change points accurately and at scale. Experiments with three configurable software systems confirm the temporal precision of our approach and have confirmed prevalent practical challenges when studying real world-software systems. It is an avenue for further research to incorporate further information, such as commit artifacts, when pinpointing performance changes.

	\section{Acknowledgments}
	Apel’s work has been supported by the Deutsche Forschungsgemeinschaft (DFG) under the contract AP 206/11-1. Siegmund’s work has been supported by the DFG under the
	contracts SI 2171/2 and SI 2171/3-1.

>>>>

	\section{Discussion}\label{sec:discussion}
	\paragraph{Efficiency} Among all experiments conducted with our synthetic setup, our approach yielded reasonable to high accuracy across most parameter settings. A subsequent analysis of parameter settings showed that our algorithm is not limited by the size of a problem space, but rather the number of measurements per iteration. 
	% mehr Interpretation, Details ausblenden
	{\color{teal}Our algorithm found multiple change points across parameter settings, including those with large numbers of commits and configuration options. While the search space of possible measurements is exponential in the number of configuration options, the theoretical maximum measurement budget before the algorithm aborted was 150,00 measurements, considering  5,000 measurements timeout of 30 iterations. The vast majority of cases terminated in early iterations with high accuracy. That is, we were able to tackle a problem with exponential complexity with a measurement budget that is linear in the number of iterations and the measurements per iteration, respectively.}
	Hence, we are confident that our approach scales and can efficiently approximate change points for large software systems. Notably, we are able to reproduce this finding in real-world settings when identifying the temporal location of change points. For our controlled experiment setup, we have not tuned the algorithms’ parameters to reach optimal efficiency or accuracy. However, even without tuning, we demonstrated the feasibility and efficiency of our approach. Project-specific fine tuning of several parameters is a possible extension to our approach and can further improve our results.

	\paragraph{Pinpointing Configuration Options}
	The synthetic setup has shown that our algorithm can conceptually handle both time and space. We were able to precisely identify commits for which performance changed substantially. By contrast, we were not able to fully reproduce this in a setting with real-world data when attempting to pinpoint change points to configuration options. {\color{teal}Possible explanations include an inconsistently documented configuration interface. Missing information in the real-world setup emphasizes the importance of a synthetic setup to evaluate approaches like ours and explore possible limitations and necessary extensions.} So, we have documented that, given a precisely identified commit, we can link the affected files to configuration options with little effort. In turn, given a set of code segments of a commit, matching this with the overall code base has been extensively studied before under the umbrella of feature location~\cite{hill_which_2013}.
	Inconsistencies in the documentation of configuration options is a well known and prevalent problem~\cite{rabkin_static_2011}.

	Our approach might be a means to complement existing feature location techniques with information about configuration-dependent performance changes. 	
	In our latter experiment, instead of directly sampling new arbitrary configurations, we used a finite set of configurations as a proxy (cf.~Section~\ref{sec:rq2_operationalization}). This does not reflect the original layout of our algorithm, but lets us efficiently explore some limitations of this approach. We have learned that, for interpreting the results of our approach, domain knowledge is advantageous as we have identified incomplete and inconsistent configuration interfaces as possible limiting factors. We acknowledge that, while our approach performs well as an automated pipeline in a controlled environment, for real-world applications, a semi-automated setup with additional manual measurement prioritization and deeper knowledge of configuration constraints would improve the applicability of our approach. This is a promising further avenue of future work.

@@

	\section{Conclusion}

	\begin{comment}
	Changes in the observed performance of a software system can be configuration-specific and difficult to pinpoint to configuration options and commits. We combine both two orthogonal perspectives, the commit history and configurability of software systems to pinpoint performance changes to individual commits and configuration options. 
	We propose an iterative sampling approach that leverages similarities in performance histories of related configurations to identify performance shifts precisely and, subsequently, pinpoint them to configuration options and interactions. 
	We have shown with a synthetic data set that our approach is sound and able to identify performance change points accurately and at scale. Experiments with three configurable software systems confirm the temporal precision of our approach and have confirmed prevalent practical challenges when studying real world-software systems. It is an avenue for further research to incorporate further information, such as commit artifacts, when pinpointing performance changes.
	\end{comment}
	{\color{teal}In this paper, we have introduced a novel aproach to identify configuration-specific performance changes across software systems‘ version histories. We conducted two sets of experiments, in a controlled setup with synthesized performance data, and in a real-world setup with performance measurements of three configurable software systems. Our experiments with synthetic data show that our approach can identify configuration-specific software performance changes accurately and at scale. 
	Our results for real-wold software systems uphold the temporal precision of our approach. We could confirm prevalent practical challenges when studying real world-software systems and we were able to pinpoint performance changes to configuration options with little additional manual effort. It is an avenue for further research to incorporate further information, such as commit artifacts, when identifying configuration-specific software performance changes automatically.}

	\section{Acknowledgments}
	We would also like to thank our reviewers for their comments. Apel’s work has been supported by the Deutsche Forschungsgemeinschaft (DFG) under the contract AP 206/11-1. Siegmund’s work has been supported by the DFG under the
	contracts SI 2171/2 and SI 2171/3-1.
